Parsing Postfix log files

### 2008-03-08 doc/logparser.tex
Diagnosed, fixed and documented another complication: mails deleted by postsuper during delivery.

Postfix logs the recipient address when delivery of a mail is attempted, so
if delivery has yet to be attempted the parser cannot determine the
recipient address or addresses.  This is a problem when mail is arriving
faster than Postfix can attempt delivery, and the administrator deletes
some of the mail (because it's the result of a mail loop, mail bomb, or
joe~job) before Postfix has had a chance to try to deliver it.  In this
case the recipient address will not have been logged, so a dummy recipient
address needs to be added, as every mail is required to have at least one
recipient.  This complication has been observed in few log files, but
typically when it does occur there will be many instances of it, closely
grouped.

@@

@@

@@

>>>>

Postfix logs the recipient address when delivery of a mail is attempted, so
if delivery has yet to be attempted the parser cannot determine the
recipient address or addresses.  This is a problem when mail is arriving
faster than Postfix can attempt delivery, and the administrator deletes
some of the mail (because it's the result of a mail loop, mail bomb, or
joe~job) before Postfix has had a chance to try to deliver it.  In this
case the recipient address will not have been logged, so a dummy recipient
address needs to be added, as every mail is required to have at least one
recipient.  This complication has been observed in few log files, but
typically when it does occur there will be many instances of it, closely
grouped.  The DELETE action is responsible for handling this complication.

@@

\subsubsection{Mails deleted during delivery}

\label{Mails deleted during delivery}

The administrator can delete mails using \daemon{postsuper}; occasionally
mails in the process of being delivered will be deleted.  This results in
the log lines from the delivery agent (\daemon{local}, \daemon{virtual} or
\daemon{smtp}) appearing in the log file \textit{after\/} the mail has been
removed from the state tables and saved in the database.  The DELETE action
adds deleted mails to a cache, which is checked by the SAVE\_BY\_QUEUEID
action, and the current log line discarded if the following conditions are
met:

    \item The queueid is not found in the state tables.

    \item The queueid is found in the cache of deleted mails.

    \item The timestamp of the log line is within 5 minutes of the end time
        of the mail.

Sadly this solution involves discarding some data, but the complication
only arises eight times (tightly grouped in one log file, which is not in
the group used for testing); if this complication occurred more frequently
it might be desirable to find the mail in the database and add the log
line to it.

@@

    \item The solution to complication \sectionref{Mails deleted during
        delivery} involves discarding data.

@@

    \item Improve the solution to complication \sectionref{Mails deleted
        during delivery} so that data is not discarded.
